### 2019-04-26 R/model.Rmd
Creating initial RF run with all data

---
title: "Untitled"
author: "Betty Kreakie"
date: "April 25, 2019"
output: html_document
---

```{r libraries, echo=TRUE, cache=TRUE}

library(randomForest)
library(dplyr)
library(purrr)
library(here)
library(broom)
library(ggplot2)
```

#Run Variable Selection stuff (hollister et al 2017)

```{r var_sel,echo=TRUE, cache=TRUE}

#varsel
source(here::here('R/varsel_regression_rf.R'))
source(here::here('R/varsel_plot.R'))

x<-complete.cases(nla)
nla.cc<-nla[x,]
nla.cc$year<-factor(nla.cc$year)

data07 <- filter(nla.cc, year == 2007, tmean_2m < 70) %>%
            dplyr::select(tmean_2m, date, longitude, latitude, tmean, tmean_dm1,
            tmean_avg3, tmean_avg7, tmean_avg30, percent_impervious, surface_area,
            shoreline_length, shoreline_dev, max_length, max_width, mean_width,
            max_depth, max_depth, volume, elevation) # 1214   19

data12 <- filter(nla.cc, year == 2012, tmean_2m < 70) %>%
            dplyr::select(tmean_2m, date, longitude, latitude, tmean, tmean_dm1,
            tmean_avg3, tmean_avg7, tmean_avg30, percent_impervious, surface_area,
            shoreline_length, shoreline_dev, max_length, max_width, mean_width,
            max_depth, max_depth, volume, elevation)

nla.all <-  select(nla.cc, tmean_2m, year, date, longitude, latitude, tmean,
                   tmean_dm1, tmean_avg3, tmean_avg7, tmean_avg30,percent_impervious,                    surface_area,shoreline_length, shoreline_dev, max_length,                            max_width, mean_width, max_depth, max_depth, volume, elevation)

var_sel_tmean07 <- varsel_regression_rf(data07$tmean_2m, select(data07, -tmean_2m),                ntree=1000, prog=TRUE)

varsel_plot(var_sel_tmean07) #9 variables

var_sel_tmean07$vars[[9]]

var_sel_tmean12 <- varsel_regression_rf(data12$tmean_2m, select(data12, -tmean_2m),                ntree=1000, prog=TRUE)

varsel_plot(var_sel_tmean12) #9 variables

var_sel_tmean12$vars[[9]]

var_sel_tmean <- varsel_regression_rf(nla.all$tmean_2m, select(nla.all, -tmean_2m),                ntree=1000, prog=TRUE)

varsel_plot(var_sel_tmean) #9 variables

var_sel_tmean$vars[[9]]

```

# Build our draft model with selected variables

```{r draft_model,echo=TRUE, cache=TRUE}

# Draft Model

# Using tmean made most sense and perfomed better that tmin and tmax based models.

RF2007 <- data07 %>%

  select(tmean_2m,var_sel_tmean07$vars[[9]]) %>%

  randomForest(tmean_2m ~ ., data = ., importance = TRUE,

               ntree = 10000)

RF2007

varImpPlot(RF2007)

RF2012 <- data12 %>%

  select(tmean_2m,var_sel_tmean12$vars[[9]]) %>%

  randomForest(tmean_2m ~ ., data = ., importance = TRUE,

               ntree = 10000)

RF2012

varImpPlot(RF2012)

RFAll <- nla.all %>%

  select(tmean_2m,var_sel_tmean$vars[[9]]) %>%

  randomForest(tmean_2m ~ ., data = ., importance = TRUE,

               ntree = 10000)

RFAll

varImpPlot(RFAll)

```

mean_run <- nla.cc %>%

  select(tmean_2m,var_sel_tmean$vars[[9]]) %>%

  randomForest(tmean_2m ~ ., data = ., importance = TRUE,

               ntree = 1000)

mean_run

varImpPlot(mean_run)

```