### 2019-04-26 data/lake_temperatures_data.Rmd

---
title: "lake_temperatures_data"
author: "bryan"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(here)
library(sf)
library(prism)
library(readr)
library(raster)
library(mapview)
library(knitr)
# library(goatscape) #use source(here::here('R/gs_nlcd.r')) instead
library(prism)
library(elevatr)
# library(goatscape)  # use: source(here::here('R/gs_nlcd.r'))
library(RSQLite)
library(readxl)

options(stringsAsFactors=FALSE)
options(prism.path = "L:/Public/Milstead_Lakes/prism")  # will be really big

# create dir "data_local" for scratch files
if(!dir.exists(here::here('data_local'))) dir.create(here::here('data_local'))

# load functions
source(here::here('R/gs_nlcd.r'))  # the goatscape gs_nlcd.r function

```

## To Do List
*

## Introduction
* The goal of this project is to compile all of the data necessary to model lake surface temperatures and make predictions for all lakes in NHDplus 
* This file replaces two preliminary attempts:
    * lake_temp_data.Rmd
    * prism_big_data.Rmd
* All data will be assembled and stored as a SQLite database (lake_temperatures.sqlite) 
    * database too large to share on github.  Must be assembled with the code below.
    * [SQLite tables and fields schema and data definitions] (https://github.com/willbmisled/photic_zone_temp/blob/master/data/lake_temperatures_sqlite.xlsx)
* This document also makes a subset of the data available as a .csv file.  This file [nla_base.csv] includes just the data for NLA lakes (2007 & 2012) that will be used to build a random forest model to predict lake temperatures.  Data definitions are at the bottom of this document.

## Data sources

* **lmorpho**: [Jeff Hollister's Lake Morphometry database](https://edg.epa.gov/data/PUBLIC/ORD/NHEERL/LakeMorphometry.zip)
* **nla**
    - [NLA2007 Lake locations shapefile](https://www.epa.gov/sites/production/files/2017-10/national_lakepoly_withmetrics_20090929.zip)
    - [NLA2012 Lake locations shapefile](https://www.epa.gov/sites/production/files/2017-10/nla_2012_sites_sampled.zip)
    - [NLA2007 design file](https://www.epa.gov/sites/production/files/2014-01/nla2007_sampledlakeinformation_20091113.csv"))
    - [NLA2012 design file ](https://www.epa.gov/sites/production/files/2016-12/nla2012_wide_siteinfo_08232016.csv)
    - [NLA2007 profile data with lake temperatures by depth](https://www.epa.gov/sites/production/files/2013-09/nla2007_profile_20091008.csv) 
    - [NLA2012 profile data with lake temperatures by depth](https://www.epa.gov/sites/production/files/2016-12/nla2012_wide_profile_08232016.csv)
* **nlcd**: [2006 & 2011 NLCD percent impervious cover in a 3000m buffer around lake](https://www.mrlc.gov/index.php)
* **prism**: [prism estimates of daily min & max temperatures for date range (sample_date - 30 days):(sample_date) by location](http://www.prism.oregonstate.edu/)

## CRS info
* prism: '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0'
* centroids: '+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0'
* lmorhpho: "+proj=merc +a=6378137 +b=6378137 +lat_ts=0.0 +lon_0=0.0 +x_0=0.0 +y_0=0 +k=1.0 +units=m +nadgrids=@null +wktext +no_defs"
* nlcd: "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0
+units=m +no_defs"

## Data steps

* Download the tmean prism daily for May1-Sep 30 for years 1981:2018

```{r get_prism, include=FALSE, echo=FALSE, eval= FALSE}
# NOte this will take many hours

for(i in 1981:2018) {
# for(i in c(1981:2006, 2008:2011, 2013:2018)) {
  min <- ymd(paste0(i, "05", "01"))
  max <- ymd(paste0(i, "09", "30"))
  print(paste(i, Sys.time()))
  get_prism_dailys(type="tmean", minDate = min, maxDate = max, keepZip=FALSE)
}}

```

* Lake Morphometry
    - download, convert to a single sf file
    - save as "lmorpho.rda"

```{r get_morpho, include=FALSE, echo=FALSE, eval=FALSE} 
#download existing lake morphometry
if(!dir.exists(here::here("data_local/LakeMorphGdb.gdb"))){
  temp <- tempfile()
  download.file("https://edg.epa.gov/data/PUBLIC/ORD/NHEERL/LakeMorphometry.zip",temp)
  unzip(temp, exdir = here::here('data_local'))
  unlink(temp)
}

# get layers
if(file.exists(here::here('data_local/lmorpho.rda'))) {
  load(here::here('data_local/lmorpho.rda'))
  } else {
    lyrs <- sf::st_layers(here::here("data_local/LakeMorphGdb.gdb"))
    lmorph <- purrr::map(lyrs$name, function(x) sf::read_sf(here::here("data_local/LakeMorphGdb.gdb/"), x))
    lmorpho <- do.call(rbind, lmorph)

  # change nlaSITE_ID == "NA" to NA
  dim(lmorpho) # 363314     19
  table(is.na(lmorpho$nlaSITE_ID)) # FALSE 363314

  lmorpho <- mutate(lmorpho, nlaSITE_ID = ifelse(nlaSITE_ID == "NA", NA, nlaSITE_ID))
  table(is.na(lmorpho$nlaSITE_ID)) # FALSE =1151  TRUE = 362163

  save(lmorpho, file = here::here('data_local/lmorpho.rda'))
}
```

* add centroids (longitude and latitude) to lmorpho
    * extract centroids from lmorpho
    * transform to prism crs (+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0)
    * add as longitude and latitude

```{r centroids, include=FALSE, echo=FALSE, eval=FALSE} 
# download the lmorpho data
load(here::here('data_local/lmorpho.rda'))

# get the centroids
centroids <- st_centroid(lmorpho)

# transform to prism CRS
# +proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0

centroids <- st_transform(centroids, "+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0")

# add lat long and convert to data.frame
xy <- as.data.frame(st_coordinates(centroids))
names(xy) <- c('longitude', 'latitude')

centroids <- as.data.frame(centroids) %>%
    dplyr::select(-Shape) %>%
    cbind(xy) 
```

* get the prism cellnums (the grid cells) for the centroids

```{r cellnums, include=FALSE, echo=FALSE, eval=FALSE} 
# get a prism raster 
pf <- raster(ls_prism_data(absPath=T)[1, 2])

# change values of grid cells to the cell number
values(pf) <- 1:ncell(pf)

# extract the cellnums from the rasters
a <- extract(x = pf, y = centroids[ , c("longitude", "latitude")])
    length(a) # 363314
    length(unique(a)) # 144402

# add cellnum to centroids
centroids$cellnum <- a

# save file
save(centroids, file = here::here('data/centroids.rda'))
```

* write a function to extract the prism data
    * each year will be separate file saved as data_local/meanYYYY.rds  with YYYY = year

```{r func_prism, include=FALSE, echo=FALSE, eval=TRUE}
# function to extract the values for all dates in a year
pvals <- function(year) {
# free up memory
  gc()
# get the dates
  dates <- format(seq(ymd(paste0(year, "05", "01")), ymd(paste0(year, "09", "30")), by="days"), format="%Y%m%d")
# get the prism file names
  pfiles <- paste0('PRISM_tmean_stable_4kmD1_', dates, '_bil')
# make a raster stack
  pstack <- prism_stack(pfiles)
# get the values for all cellnums
  vals <- as.data.frame(getValues(pstack))
# shorten names to dmmdd (e.g., d0501)
  names(vals) <- paste0("d", substr(dates, 5, 8))
# add cellnum (same as the rowname)
  vals$cellnum <- as.numeric(rownames(vals)) 
# add year
  vals$year <- year
# keep only cellnums (cnums) with lakes in them
  vals <- left_join(cnums, vals) 
# assign df names as meanYYYY (year from input)
  fname <- paste0('mean', year)
  assign(fname, vals)
# save the file
  outname <- paste0('data_local/', fname, '.rds')
  saveRDS(object = get(fname), file = here::here(outname)) 
}

```

* Extract the values from the prism rasters and add to the table tmeans_raw in lake_temperatures.sqlite

```{r vals_prism, include=FALSE, echo=FALSE, eval=FALSE}
# create a list of cellnums is available
if(exists("cnums") == FALSE){
  if(exists("centroids") == FALSE) load(here::here('data/centroids.rda'))
  cnums <- data.frame(cellnum = unique(centroids$cellnum)) %>% # 144402      1
    arrange(cellnum)
}

# loop to extract values (note a file for each year will be saved to /data/ but not stored in memory)
for (i in c(1981:2017)) {
  if(!file.exists(paste0('data_local/mean', i, '.rds'))) {
    print(i)
    print(Sys.time())
    pvals(i)
}}

# combine all means into a single SQLite table

# create or connect to the database
  means <- dbConnect(RSQLite::SQLite(), here::here("data_local/lake_temperatures.sqlite"))

# append data to table means
  for(i in c(1981:2017)) {
  print(Sys.time())
  print(i)
  rds <- paste0('data_local/mean', i, '.rds')
  a <- readRDS(here::here(rds))
  dbWriteTable(means, "means", a, append = TRUE)
  print(memory.size())
  }

# let's rename table 'means' to 'tmeans_raw'
dbListTables(means)
dbExecute(means, 'ALTER TABLE means RENAME TO tmeans_raw')
dbListFields(means, "tmeans_raw")

# create indices for cellnum and year for table "tmeans_raw"
means <- dbConnect(RSQLite::SQLite(), here::here("data_local/lake_temperatures.sqlite"))
dbExecute(means,"CREATE INDEX index_tmeans_raw_cnum ON tmeans_raw (cellnum)")
dbExecute(means,"CREATE INDEX index_tmeans_raw_year ON tmeans_raw (year)")
dbDisconnect(means)

# close and save the database
  dbDisconnect(means)
```

* calculate 3, 7, and 30 mean temps
    * means calculated and added to the table tmeans in lake_temperatures.sqlite
    * this takes a long time to eval = FALSE; change if you want to rerun

```{r means_prism, include=FALSE, echo=FALSE, eval=FALSE}

for(j in c(1983:2017)) { 
# for(j in c(1981:1982)) {   
  print(Sys.time())
# load a year of data
  start <- Sys.time()
  print(j)
  rds <- paste0('data_local/mean', j, '.rds')
  tmean <- readRDS(here::here(rds))

# create dfs to hold the preliminary results
  tmean3 <- tmean[ , -c(2:32)] 
  tmean3[ , 2:123] <- NA
  tmean7 <- tmean3
  tmean30 <- tmean3

# create df of temps for day-1
  tmean_dm1 <- tmean3
  tmean_dm1[ ,2:123] <- tmean[ ,32:153]

# loop to calculate the 3, 7, and 30 day means 
for(i in c(33:154)){
  tmean3[ , i-31] <- apply(tmean[ ,(i-3):(i-1)], 1, mean)
  tmean7[ , i-31] <- apply(tmean[ ,(i-7):(i-1)], 1, mean)
  tmean30[ , i-31] <- apply(tmean[ ,(i-30):(i-1)], 1, mean)
}

# convert to long format
  tm <- gather(tmean[ , -c(2:32)], date, tmean, -cellnum, -year)
  tm_dm1 <- gather(tmean_dm1, date, tmean_dm1, -cellnum, -year)
  tm3 <- gather(tmean3, date, tmean, -cellnum, -year)
  tm7 <- gather(tmean7, date, tmean, -cellnum, -year)
  tm30 <- gather(tmean30, date, tmean, -cellnum, -year)

# combine year and date
  tm <- dplyr::mutate(tm, date = paste0(tm$year, substr(tm$date, 2, 5))) %>%
    dplyr::select(-year)

# add dm1 and 3, 7, and 30 day means
  tm$tmean_dm1 <- tm_dm1$tmean_dm1
  tm$tmean_avg3 <- tm3$tmean
  tm$tmean_avg7 <- tm7$tmean
  tm$tmean_avg30 <- tm30$tmean

# connect to the database
  means <- dbConnect(RSQLite::SQLite(), here::here("data_local/lake_temperatures.sqlite"))

# append data to table tmeans
  dbWriteTable(means, "tmeans", tm, append = TRUE)

# QAQC
  # dbListTables(means)
  # dbGetQuery(means, 'SELECT Count(*) FROM tmeans') # 17617044 (after first upload)
  # dim(tm) # 17617044        6

# close and save the database
  dbDisconnect(means)

print(Sys.time() - start)
}

# create indices for cellnum and date for table "tmeans"
means <- dbConnect(RSQLite::SQLite(), here::here("data_local/lake_temperatures.sqlite"))
dbExecute(means,"CREATE INDEX index_tmeans_cnum ON tmeans (cellnum)") 
dbExecute(means,"CREATE INDEX index_tmeans_date ON tmeans (date)")
dbDisconnect(means)

# QAQC
# check number of lines
means <- dbConnect(RSQLite::SQLite(), here::here("data_local/lake_temperatures.sqlite"))
dbGetQuery(means, 'SELECT Count(*) FROM tmeans') # 17617044 * 37 = 651,830,628

# check some of the means for a random cellnums and dates
means <- dbConnect(RSQLite::SQLite(), here::here("data_local/lake_temperatures.sqlite"))

yr1 <- sample(1981:2017, 1)
day1 <- sample(format(seq(ymd(paste0(yr1, "06", "01")), ymd(paste0(yr1, "09", "30")), by="days"), format="%m%d"), 1)
date1 <- paste0(yr1, day1)

mean1 <- dbGetQuery(means, paste0('SELECT * FROM tmeans WHERE cellnum = ', cn, ' AND date = ', date1))

mean2 <- dbGetQuery(means, paste0('SELECT * FROM tmeans_raw WHERE cellnum = ', cn, ' AND year = ', yr1))

dbDisconnect(means)

i <- which(names(mean2) == paste0('d', day1))
c(mean1$tmean == mean2[i],
mean1$avg_3day_tmean == apply(mean2[ ,(i-2):i], 1, mean),
mean1$avg_7day_tmean == apply(mean2[ ,(i-6):i], 1, mean),
mean1$avg_30day_tmean == apply(mean2[ ,(i-29):i], 1, mean))
##

#

# qa qc repeat in excel for one year (use a subset of the cellnums) - results are the same
  # sample 100 cellnums from mean1981
mean2001 <- readRDS(here::here('data_local/mean2001.rds'))
temp <- mean2001[sample(as.numeric(row.names(mean2001)), 100), ] %>%
      arrange(cellnum)

  #get the calculated means for the sampled cellnums
means <- dbConnect(RSQLite::SQLite(), here::here("data_local/lake_temperatures.sqlite"))

aves <- dbGetQuery(means, paste0('SELECT * FROM tmeans WHERE date IN (20010601, 20010707, 20010930) AND
                                 cellnum IN (', paste(temp$cellnum, collapse=", "), ')')) %>%
        arrange(cellnum)
dbDisconnect(means)

  # write data to csv
write.csv(temp, here::here('data_local/temp/tmean.csv'))
write.csv(aves, here::here('data_local/temp/tmean_avg.csv'))

  # calculate the means in excel and compare to the results above
excel <- read.csv(here::here('data_local/temp/tmean_excel.csv'))
all.equal(aves$cellnum, excel$cellnum) # TRUE
all.equal(as.numeric(aves$date), excel$date) # TRUE
all.equal(aves$tmean, excel$tmean) # TRUE
all.equal(aves$tmean_dm1, excel$tmean_dm1) # TRUE
all.equal(aves$tmean_avg3, excel$tmean_avg3) # TRUE
all.equal(aves$tmean_avg7, excel$tmean_avg7) # TRUE
all.equal(aves$tmean_avg30, excel$tmean_avg30) # TRUE

```

* get the elevations for the centroids

```{r elevatr, include=FALSE, echo=FALSE, eval=FALSE}
locations <- dplyr::select(centroids, longitude, latitude, COMID)

prj <- "+proj=longlat +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +no_defs"
start <- Sys.time()
elev <- get_elev_point(locations, prj = prj, src = 'aws', z = 8)
end <- Sys.time()
end - start # 36.65486 mins

dim(centroids) # 363314     21
dim(elev) # 363314      3
centroids <- left_join(centroids, elev@data)
dim(centroids) # 363328     23

a <- as.data.frame(table(elev$COMID))
b <- dplyr::filter(a, Freq > 1)

dplyr::filter(elev@data, COMID %in% b$Var1)

# Why are there 14 extra points now?
length(unique(centroids$COMID)) # 363307  This includes the seve

saveRDS(object = elev, file = here::here('data/elevations.Rds'))

```

* create the lakes table from lmorpho, centroids, and elevation
* seven COMIDs removed because they were repeated and had slighly different data
* select and rename fields

```{r lakes, include=FALSE, echo=FALSE, eval=FALSE}
# check centroids / morpho for repeated comids
a <- as.data.frame(table(centroids$COMID))
b <- dplyr::filter(a, Freq > 1)  # 7 comids are repeated

lakes <- dplyr::filter(centroids, !COMID %in% b$Var1)
dim(centroids) # 363314     21
dim(lakes) # 363300     21; removed 7 centroids repeated 2 x each = 14 rows

# add elevation data
lakes <- left_join(lakes, elev@data) # 363300     23

# select and rename
lakes <- dplyr::select(lakes, cellnum, comid = COMID, longitude, latitude, 
                        surface_area = SurfaceArea,
                        shoreline_length = ShorelineLength,
                        shoreline_dev = ShorelineDev,
                        max_length = MaxLength, 
                        max_width = MaxWidth,
                        mean_width = MeanWidth,
                        max_depth = MaxDepthCorrect,
                        mean_depth = MeanDepthCorrect,
                        volume = VolumeCorrect,
                        elevation)

# create or connect to the database
  means <- dbConnect(RSQLite::SQLite(), here::here("data_local/lake_temperatures.sqlite"))

# add lakes to table lakes
  dbWriteTable(means, "lakes", lakes, overwrite = TRUE)

# index table lakes
  dbExecute(means,"CREATE INDEX index_lakes_cnum ON lakes (cellnum)") 
  dbExecute(means,"CREATE INDEX index_lakes_comid ON lakes (comid)")

# housekeeping  
dbListTables(means)
dbListFields(means, "lakes")

# close and save the database
  dbDisconnect(means)

# save an Rds copy of lakes
  saveRDS(object = lakes, file = here::here('data/elevations.Rds'))

```

* NLA 
    - download the nla2007 & nla2012 shapefiles
    - transform to st_crs(lmorpho) 
        -  EPSG: 3857
        -  proj4string: "+proj=merc +a=6378137 +b=6378137 +lat_ts=0.0 +lon_0=0.0 +x_0=0.0 +y_0=0 +k=1.0 +units=m +nadgrids=@null +wktext +no_defs"
    - use st_join to match the 2007 and 2012 lakes to lmorpho
      - 54 nla07 lakes removed due to lack of 1:1 correspondence with lmorpho
      - 7 nla12 lakes removed because they did not map to any lmorpho lakes
      - 82 nla12 lakes removed due to lack of 1:1 correspondence with lmorpho
    - rbind nla2007 and nla2012

```{r get_nla_gis, include=FALSE, echo=FALSE, eval=FALSE}

# download nla07
  if(!file.exists(here::here('data_local/National_LakePoly_withMetrics_20090929.shp'))) {
    temp <- tempfile()
    download.file("https://www.epa.gov/sites/production/files/2017-10/national_lakepoly_withmetrics_20090929.zip",temp)
    unzip(temp, exdir = here::here('data_local'))
    unlink(temp)
  }

# download nla12
  if(!file.exists(here::here('data_local/NLA2012_Sampled_Polys_20131217.shp'))) {
    temp <- tempfile()
    download.file("https://www.epa.gov/sites/production/files/2017-10/nla_2012_sites_sampled.zip",temp)
    unzip(temp, exdir = here::here('data_local'))
    unlink(temp)
  }

# transform NLA shapefiles to st_crs(lmorpho)
  nla07 <- st_read(here::here('data_local/National_LakePoly_withMetrics_20090929.shp')) %>%
            st_transform(st_crs(lmorpho))  # dim = 1159    6

  nla12 <- st_read(here::here('data_local/NLA2012_Sampled_Polys_20131217.shp')) %>%
            st_transform(st_crs(lmorpho))  # dim = 1269 153

# join nla to lmorpho
  j07 <- st_join(nla07, lmorpho) %>%
          dplyr::select(nla_id = SITEID, comid = COMID) %>%
          mutate(year = 2007)

  j12 <- st_join(nla12, lmorpho) %>%
          dplyr::select(nla_id = NLA12_ID, comid = COMID.y) %>%
          mutate(year = 2012)

# convert joins to data.frames
  st_geometry(j07) <- NULL
  st_geometry(j12) <- NULL

# remove repeated comid's and and nla_ids; lakes without a 1:1 correspondence
#2007
  dim(nla07) # 1159    6
  dim(j07) # 1186   3
  table(is.na(j07$comid)) # FALSE = 1186
  table(is.na(j07$nla_id)) # FALSE = 1186

  a <- as.data.frame(table(j07$comid))
  bc <- filter(a, Freq > 1) # 4 lmorpho lakes map to multiple nla lakes map; delete these

  a <- as.data.frame(table(j07$nla_id))
  bs <- filter(a, Freq > 1) # 18 nla lakes map to multiple lmorpho lakes map; delete these

  j07 <- filter(j07, !comid %in% bc$Var1, !nla_id %in% bs$Var1) # 1132    3

#2012
  dim(nla12) # 1269  153
  dim(j12) # 1315    3
  table(is.na(j12$comid)) # FALSE = 1308 TRUE = 7
  table(is.na(j12$nla_id)) # FALSE = 1315

  j12 <- j12[!is.na(j12$comid), ] # 1308    3

  a <- as.data.frame(table(j12$comid))
  bc <- filter(a, Freq > 1) # 2 lmorpho lakes map to multiple nla lakes map; delete these

  a <- as.data.frame(table(j12$nla_id))
  bs <- filter(a, Freq > 1) # 32 nla lakes map to multiple lmorpho lakes map; delete these

  j12 <- filter(j12, !comid %in% bc$Var1, !nla_id %in% bs$Var1) # 1226    3

# rbind
  nla_id <- rbind(j07, j12) # 2358    3

# save an Rds copy of lakes
  saveRDS(object = nla_id, file = here::here('data/nla_id.Rds')) 
  # nla_id <- readRDS(file = here::here('data/nla_id.Rds'))
```

* get the nla profile data
* calc the mean temperature for top 2m by nla_id, date, and visit_no
* join to nla_id to create table nla
* no profile data for 12 nla07 and 160 nla12 lakes
* 184 lakes sampled twice
* 4 lakes sampled on visit_no == 2 only

```{r nla_data, include=FALSE, echo=FALSE, eval=FALSE} 
# get the profile data
  p07 <- read.csv(url("https://www.epa.gov/sites/production/files/2013-09/nla2007_profile_20091008.csv")) %>% 
    dplyr::select(nla_id = SITE_ID, visit_no = VISIT_NO, date = DATE_PROFILE, depth = DEPTH, 
                  temperature = TEMP_FIELD) %>% 
    dplyr::filter( !is.na(temperature)) %>%
    dplyr::mutate(date = format(mdy(date), format="%Y%m%d")) # 12648     5

   p12 <- read.csv(url("https://www.epa.gov/sites/production/files/2016-12/nla2012_wide_profile_08232016.csv")) %>%
    filter(SAMPLE_TYPE == "PROF", SITE_ID != '', !is.na(DEPTH), !is.na(TEMPERATURE)) %>%
    dplyr::select(nla_id = SITE_ID, visit_no = VISIT_NO, date = DATE_COL, depth = DEPTH, 
                  temperature = TEMPERATURE) %>%
    dplyr::mutate(date = format(dmy(date), format="%Y%m%d")) # 10967     5

#rbind p07 & p12  
  prof <- rbind(p07, p12)
  p2m <- dplyr::filter(prof, depth <= 2) # 9045    5

  tmean_2m <- group_by(p2m, nla_id, visit_no, date) %>% 
    summarise(tmean_2m = mean(temperature), tmean_2m_n = n()) %>%
    left_join(nla_id) %>%
    dplyr::filter(!is.na(comid)) %>%
    dplyr::select(nla_id, comid, year, date, visit_no, tmean_2m, tmean_2m_n) %>%
    ungroup()

# some lakes have missing profile data  
  dim(tmean_2m) # 2370    7
  length(unique(tmean_2m$nla_id)) # 2186
  a <- filter(nla_id, !nla_id %in% tmean_2m$nla_id)
  table(a$year) # 2007 = 12; 2012 = 160

# some lakes sampled twice
  table(tmean_2m$visit_no) # 1 = 2182; 2 = 188

  a <- as.data.frame(table(tmean_2m$nla_id))
  b <- filter(a, Freq > 1) # 184 2; so 4 lakes sampled on visit_no == 2 only

# save tmean_2m
        saveRDS(object = tmean_2m, file = here::here('data/tmean_2m.Rds'))
```

### nlcd impervious data
* calculate impervious surface in a 3000m buffer  for lmorpho / NLA lakes
    - use year = 2006 for the 2007 lakes
    - use year = 2011 for the 2012 lakes
* restrict to lakes in tmean_2m since these have water temperature data.
* 3k buffer also used in the ecosphere paper: https://esajournals.onlinelibrary.wiley.com/doi/epdf/10.1002/ecs2.1321
* The buffering failed for comid == 14489102 (nla_id == NLA06608-0243 & nla_id == NLA12_SD-101); for now this lake will be removed.
* goatscape did not return data for 14 of the 2007 lakes and 2 of the 2012 lakes.  We could work on this further but it is easiest just to eliminate these.  
* The percent overlap field shows what percent of the buffer is covered by NLCD.  For 2007 and 2012 1 and 2 lakes are not completely covered respectively.  Eliminate these. 
* For QAQC percent impervious was calculated in ArcMap for a couple of lakes.  Results comparable.

```{r imperv, include=FALSE, echo=FALSE, eval=FALSE} 
if(!exists("tmean_2m")) tmean_2m <- readRDS(here::here('data/tmean_2m.Rds'))
if(!exists("lmorpho")) load(here::here('data_local/lmorpho.rda'))

#  albers projection used by nlcd
crs_alb <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +ellps=GRS80 
  +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
   landscape<-sf::st_transform(landscape, crs_alb) #reproject to the Albers Equal Area Conic projection used by NLCD

# get the comids for nla lakes from tmean_2m
# remove comid == 14489102; it fails in the buffer step
  ids <- dplyr::select(tmean_2m, nla_id, comid, year) %>%   # 2370    3
          filter(comid != 14489102) %>%                     # 2367    3
          distinct()                                        # 2184    3

  filter(ids, year == 2007) %>% dplyr::select(comid) %>% distinct() %>% dim() # 1119    1
  filter(ids, year == 2012) %>% dplyr::select(comid) %>% distinct() %>% dim() # 1065    1; 1119 + 1065 = 2184

# get the % imperv
# set up data.frame for results
impervious <- data.frame(comid = as.numeric(rep(NA, length(ids$comid))),
                    nla_id = as.character(rep(NA, length(ids$comid))),
                    total_percent_impervious = as.numeric(rep(NA, length(ids$comid))),
                    total_areaM2_impervious = as.numeric(rep(NA, length(ids$comid))),
                    buf_area_m2 = as.numeric(rep(NA, length(ids$comid))),
                    percent_overlap = as.numeric(rep(NA, length(ids$comid))))

# loop to populate data.frame
for(i in min(which(is.na(impervious))):length(ids$comid)) {
  #for(i in min(which(is.na(impervious))):5) {
start <- Sys.time()
lago <- tryCatch(st_transform(filter(lmorpho, comid == ids$comid[i]), crs_alb), error = function(e) NA) 
buf <- tryCatch(sf::st_difference(sf::st_buffer(lago, 3000), sf::st_geometry(lago)), error = function(e) NA) 
imp <- tryCatch(gs_nlcd(buf, spatial = TRUE, year = ids$year[i]-1, dataset = "impervious"), 
                error = function(e) NA)

# harvest data
impervious$comid[i] <- ids$comid[i]
impervious$nla_id[i] <- ids$nla_id[i]

if(!is.na(imp)[1]){
  impervious$total_percent_impervious[i] <- imp$total_percent_impervious
  impervious$total_areaM2_impervious[i] <- imp$total_areaM2_impervious
  impervious$buf_area_m2[i] <- sum(imp$nlcd_freq$areaM2)
  impervious$percent_overlap[i] <- imp$percent_overlap
}

plot(buf[1], main = paste0("i = ", i, " of ", nrow(ids), "; nla_id = ", ids$nla_id[i], "; end = ", 
                           format(Sys.time(), "%H:%M:%OS"))) 
print(i)
print(Sys.time()-start)
}

# remove observations where is.na(total_percent_impervious) or percent_overlap < 100
dim(impervious) # 2184    6
impervious <- filter(impervious, !is.na(impervious$total_percent_impervious)) # 2168    6
impervious <- filter(impervious, percent_overlap == 100) #  2168    6

# save impervious
    saveRDS(object = impervious, file = here::here('data/impervious.Rds'))
    # impervious <- readRDS(file = here::here('data/impervious.Rds'))

```

* create table "nla"
    * primary keys are nla_id & visit_no

```{r tbl_nla, include=FALSE, echo=FALSE, eval=FALSE} 
tmean_2m <- readRDS(file = here::here('data/tmean_2m.Rds')) # 2370 7
# tmean_2m$date <- as.numeric(tmean_2m$date)

impervious <- readRDS(file = here::here('data/impervious.Rds')) # 2168    6
imp <- dplyr::select(impervious, nla_id, comid, 
                                         percent_impervious = total_percent_impervious, 
                                         area_impervious = total_areaM2_impervious, area_total = buf_area_m2)

nla <- left_join(tmean_2m, imp) # 2370    10
  dim(unique(dplyr::select(nla, nla_id, visit_no))) # 2370 2

# connect to the database
  means <- dbConnect(RSQLite::SQLite(), here::here("data_local/lake_temperatures.sqlite"))

# add nla to table nla
  dbWriteTable(means, "nla", nla, overwrite = TRUE)

# index table nla
  dbExecute(means,"CREATE INDEX index_nla_comid ON nla (comid)") 
      #  dbExecute(means,"DROP INDEX index_nla_comid")
  dbExecute(means,"CREATE INDEX index_nlaid_comid ON nla (nla_id)")

# close and save the database
  dbDisconnect(means)
```

### QAQC tmeans_raw: checked and ready to go
* Use arcMap to extract the prism values for 1 date and compare to table tmean_raw
    * open arcMap and add prism raster tmeans for 19920930 ('data_local/prism_qaqc.mxd')
    * save table "lakes" as lakes.csv
    * import into arcmap and save as lakes.shp
    * use the extract multi point tools to add prism tmean to lakes.shp (.dbf)
    * bring back into r and compare
    * everything peachy: 
      * median difference = 0.0000006
      * max difference = 2.7 # checked this out and the lake is on the border of two grid cells
* no data for 16 cellnums corresponding to 22 lakes 
    * looked at these in arcmap and all are just outside the prism raster; no data

```{r tm_raw, include=FALSE, echo=FALSE, eval=FALSE} 
# create a shapefile from table "lakes"
means <- dbConnect(RSQLite::SQLite(), here::here("data_local/lake_temperatures.sqlite"))
dbListFields(means, "tmeans_raw")

lakes <- dbGetQuery(means, 'SELECT lakes.* FROM "lakes"') # 363300     14
  write.csv(lakes, file = here::here('data_local/lakes.csv'), row.names = FALSE)

# import into arcmap and save as shapefile 'data_local/lakes.shp'
# compare to tmeans_raw
am <- read.csv(here::here('data_local/lakes.dbf.csv')) # 363300      5
db <- dbGetQuery(means, 'SELECT cellnum, d0930 FROM "tmeans_raw" WHERE year = 1992') # 144402      2
comp <- left_join(am, db) # 363300      6
comp <- comp[complete.cases(comp), ] # 363278      6
comp$dif <- abs(comp$d0930 - comp$tm19920930)
comp <- arrange(comp, dif)
summary(comp$dif)

plot(comp$d0930, comp$tm19920930)

# missing values
table(is.na(db$cellnum)) # FALSE = 144402
table(is.na(db$d0930)) # FALSE = 144386 TRUE = 16

chk <- db$cellnum[which(is.na(db$d0930))]

# close and save the database
dbDisconnect(means)
```

### QAQC tmeans
* A date chosen randomly
* Get the raw prism data from table tmean_raw for the date and the 30 days prior for all cellnums
* Re-calculate the values in table tmeans from the raw data for all cellnums
* Compare the results to table tmeans
    * Done several times - all values match
* Re-extract the values for the date from the prism raster and compare
    * Also done several times - all values match
    * This also verifies the cellnum field in table 'lakes'

```{r tm, include=FALSE, echo=FALSE, eval=FALSE} 
# choose a random date
means <- dbConnect(RSQLite::SQLite(), here::here("data_local/lake_temperatures.sqlite"))
dbListFields(means, "tmeans")

# list of all dates
dates <- dbGetQuery(means, 'SELECT DISTINCT date FROM "tmeans"') # 4514    1

# choose a date at random 
d <- sample(dates$date, 1)
yr <- substr(d, 1, 4)
fields <- c('cellnum', 'year', paste0("d", substr(format(seq(ymd(d)-30, ymd(d), by="days"), format="%Y%m%d"), 5, 8)))

# get tmean_raw data for date
tmr <- dbGetQuery(means, paste("SELECT * FROM 'tmeans_raw' WHERE year = ", yr))
tmr <- tmr[ , fields] # 144402     33

comp <- dplyr::select(tmr, cellnum)
comp$date <- d
comp$tmean <- tmr[ , fields[33]]
comp$tmean_dm1 <- tmr[ , fields[32]]
comp$tmean_avg3 <- apply(tmr[, 30:32], 1, mean)
comp$tmean_avg7 <- apply(tmr[, 26:32], 1, mean)
comp$tmean_avg30 <- apply(tmr[, 3:32], 1, mean)

aves <- dbGetQuery(means, paste0("SELECT * FROM 'tmeans' WHERE date = ", d)) # 144402      7

all.equal(comp$cellnum, aves$cellnum) # TRUE
all.equal(comp$tmean, aves$tmean) # TRUE
all.equal(comp$tmean_dm1, aves$tmean_dm1) # TRUE
all.equal(comp$tmean_avg3, aves$tmean_avg3) # TRUE
all.equal(comp$tmean_avg7, aves$tmean_avg7) # TRUE
all.equal(comp$tmean_avg30, aves$tmean_avg30) # TRUE

# check that the cellnum field in table "lakes" is correct
# get the tmeans_raw data for the date
tmrd <- dplyr::select(tmr, cellnum, paste0('d', substr(d, 5, 8)))

# get the prism tmeans raster values for the date
pr <- raster(grep(paste0("PRISM_tmean_stable_4kmD1_", d), ls_prism_data(absPath=T)[,2],value=T))
prv <- data.frame(value = values(pr))
prv$cellnum <- as.numeric(row.names(prv))

# join and compare
chk <- left_join(tmrd, prv)
all.equal(chk[ ,2], chk[ ,3]) # TRUE

# close and save the database
dbDisconnect(means)
```

### QAQc impervious
* repeat the impervious buffer clip in arcmap
* join the nla to lmorpho and transform to nlcd crs
* write nla lakes to 'data_local/nla.shp'
* download 2011 % Dev. Imperv. CONUS: (https://www.mrlc.gov/data?f%5B0%5D=category%3Aurban%20imperviousness)
* add nla.shp and NLCD imperv to arcmap ('data_local/nlcd_qaqc.mxd)
* select nla 2012 lakes for visit_no == 2 
* buffer with "buffer" tool 
    * Distance = linear unit = 3000m
    * Side type = Outside Only
    * End Type = Round
    * Method = Planar
    * Dissolve Type = None
    * save as: nla2012_visit2_buf3000
* use the clip (Data Management) tool to clip 2011 nlcd raster to the buffer 
    * input raster = nlcd imp 2011
    * output extent = nla2012_visit2_buf3000
    * Check Box: Use input Features for Clipping Geometry = Checked
    * NoData = 255
    * save as: nla12_v1_buf3000_imp.tif
* use Zonal Statistics as Table to sumarize raster clip
    * input ... zone data = nla2012_visit2_buf3000
    * zone field = 'nla_id'
    * nla12_v1_buf3000_imp.tif
    * output table = nla12_v1_buf3000_imp.dbf
* compare arcmap and nla.impervious values graphically - near perfect fit.  Small differences expected.

```{r chk_imp, include=FALSE, echo=FALSE, eval=FALSE} 
# get the nlcd crs
a <- raster::raster(here::here('nlcd_data/cropped/buf_NLCD_2006_impervious.tif'))
crs_nlcd <- as.character(raster::crs(a))

# join nla and lmorpho
load(here::here('data_local/lmorpho.rda'))
  names(lmorpho)[1] <- 'comid'
means <- dbConnect(RSQLite::SQLite(), here::here("data_local/lake_temperatures.sqlite"))
  nla <- dbGetQuery(means, 'SELECT nla.* FROM "nla"') # 2370    9
        dbDisconnect(means)

names_nla <- names(nla)

nla <- left_join(lmorpho, nla) %>%
    dplyr::filter(!is.na(nla_id)) %>%
    dplyr::select(names_nla) %>% # 2370   10
    sf::st_transform(crs_nlcd)

# st_write(nla, here::here('data_local/nla.shp'), delete_layer = TRUE)

# join nla & arcmap summary nla12_v1_buf3000_imp.dbf
am12 <- foreign::read.dbf(here::here("data_local/nla12_v1_buf3000_imp.dbf"))
chk_am12 <- left_join(am12, nla) %>% 
          dplyr::select(nla_id, arc_per_imp = MEAN, percent_impervious, arc_area = AREA, area_total)

# compare by plotting
plot(chk_am12$arc_per_imp, chk_am12$percent_impervious)
  abline(0, 1)

plot(chk_am12$arc_area, chk_am12$area_total)
  abline(0, 1)
```

### QAQC leftovers
* The database elevation compared to the nla design file elevations; excellent match
* tmean_n recalculated in excel and compared to database; all.equal == TRUE

```{r chk_nla, include=FALSE, echo=FALSE, eval=FALSE} 
######### elevation
# get the the [elevatr] elevations
means <- dbConnect(RSQLite::SQLite(), here::here("data_local/lake_temperatures.sqlite"))
  nla <- dbGetQuery(means, 'SELECT nla.nla_id, lakes.elevation, tmean_2m, tmean_2m_n 
                          FROM "nla" 
                          INNER JOIN "lakes"
                          ON nla.comid = lakes.comid
                          WHERE nla.visit_no = 1') # 2182    4
              dbDisconnect(means)

# get the lake elevations from the design files
n07 <- read.csv(
    url("https://www.epa.gov/sites/production/files/2014-01/nla2007_sampledlakeinformation_20091113.csv"))

n07a <- filter(n07, VISIT_NO == 1) %>% 
          dplyr::select(nla_id = SITE_ID, nla_elev = ELEV_PT)

n12 <- read.csv(
    url("https://www.epa.gov/sites/production/files/2016-12/nla2012_wide_siteinfo_08232016.csv"))

n12a <- filter(n12, VISIT_NO == 1) %>% 
          dplyr::select(nla_id = SITE_ID, nla_elev = ELEVATION)

nla_elev <- rbind(n07a, n12a)

# join elevation files
elev <- left_join(nla, nla_elev)

# plot to compare
plot(elev$elevation, elev$nla_elev)
  abline(0, 1, col = "red", lwd =3)

######## tmean_2m

# get the profile data
  p07 <- read.csv(url("https://www.epa.gov/sites/production/files/2013-09/nla2007_profile_20091008.csv")) %>% 
    dplyr::select(nla_id = SITE_ID, visit_no = VISIT_NO, date = DATE_PROFILE, depth = DEPTH, 
                  temperature = TEMP_FIELD) %>% 
    dplyr::filter( !is.na(temperature)) %>%
    dplyr::mutate(date = format(mdy(date), format="%Y%m%d")) # 12648     5

   p12 <- read.csv(url("https://www.epa.gov/sites/production/files/2016-12/nla2012_wide_profile_08232016.csv")) %>%
    filter(SAMPLE_TYPE == "PROF", SITE_ID != '', !is.na(DEPTH), !is.na(TEMPERATURE)) %>%
    dplyr::select(nla_id = SITE_ID, visit_no = VISIT_NO, date = DATE_COL, depth = DEPTH, 
                  temperature = TEMPERATURE) %>%
    dplyr::mutate(date = format(dmy(date), format="%Y%m%d")) # 10967     5

# rbind p07 & p12  
  prof <- rbind(p07, p12)
  p2m <- dplyr::filter(prof, depth <= 2, visit_no == 1) # 8322    5

# write to csv
  write.csv(p2m, here::here("data_local/ptm.csv"), row.names = FALSE)

# used excel pivot table to recalculate the means.  
# compare to nla

excel <- read.csv(here::here("data_local/ptm.csv")) %>%
  dplyr::select(nla_id = nla_id_sum, temp_mean, temp_n)

tcomp <- inner_join(nla, excel)
all.equal(tcomp$tmean_2m, tcomp$temp_mean)  # TRUE
all.equal(tcomp$tmean_2m_n, tcomp$temp_n)  # TRUE
```

## Document database
* create a table in the database of data definitions (data_def) from 'data/lake_temperatures.xlsx'

```{r doc_db, include=FALSE, echo=FALSE, eval=TRUE, cache=TRUE} 
dd <- read_xlsx(here::here('data/lake_temperatures_sqlite.xlsx'), sheet = "data_defs") %>%
  dplyr::select(-rmd)

# add table data_def to the database
means <- dbConnect(RSQLite::SQLite(), here::here("data_local/lake_temperatures.sqlite"))
dbWriteTable(means, "data_defs", dd, overwrite = TRUE)

tables <- dbListTables(means)
tables

# data_defs
  dd_f <- dbListFields(means, "data_defs")  # fields
  dd_n <- dbGetQuery(means, 'SELECT Count(*) FROM data_defs') # 46 observations

# lakes
  l_f <- dbListFields(means, "lakes")  # fields
  l_n <- dbGetQuery(means, 'SELECT Count(*) FROM lakes') # 363300 observations
  dim(dbGetQuery(means, 'SELECT DISTINCT comid FROM lakes')) # 363300; primary key

# nla
  n_f <- dbListFields(means, "nla")  # fields
  n_n <- dbGetQuery(means, 'SELECT Count(*) FROM nla') # 2370 observations
  dim(dbGetQuery(means, 'SELECT DISTINCT nla_id, visit_no FROM nla')) # 2370; composite primary key

# tmeans
  tm_f <- dbListFields(means, "tmeans")  # fields
  tm_n <- dbGetQuery(means, 'SELECT Count(*) FROM tmeans') # 651830628 (651.8 m) observations
  dim(dbGetQuery(means, 'SELECT DISTINCT cellnum FROM tmeans')) # 144402; composite primary key
  dim(dbGetQuery(means, 'SELECT DISTINCT date FROM tmeans')) # 4514; composite primary key
    # NOTE:  144402 * 4514 = 651830628

# tmeans_raw
  tmr_f <- dbListFields(means, "tmeans_raw")  # fields
  tmr_n <- dbGetQuery(means, 'SELECT Count(*) FROM tmeans_raw') # 5342874 (5.3 m) observations
  dim(dbGetQuery(means, 'SELECT DISTINCT cellnum FROM tmeans_raw')) # 144402; composite primary key
  dim(dbGetQuery(means, 'SELECT DISTINCT year FROM tmeans_raw')) # 37; composite primary key
    # NOTE:  144402 * 37 = 5342874

# close and save the database
dbDisconnect(means)

# database size
size <- paste(round(file.size(here::here('data_local/lake_temperatures.sqlite')) / 1073741824, 1), 'gb')
```

* The database ('data_local/lake_temperatures.sqlite') is too large (`r size`) to keep on github.  It will have to be compiled by running the code chunks in this document.  It could be a lengthy and difficult task.

* The database includes the `r length(tables)` tables: `r tables`

#### Data definitions for table "data_defs"  
* This table has the data definitions for all tables in [lake_temperatures.sqlite]
* observations = `r dd_n` 
* fields = `r dd_f`

field	| 	description
----------- | --------------------------------- 
**table**|database table for data definitions
**field**|field in database table for data definitions
**source**|source of the data for the field
**units**|units of measurement
**indexed**|does a database index exist for the field
**primary**|Is this field a primary field?  NA = no; primary = simple primary key; composite = primary key composed of 2 or more fields
**description**|a general description of the field

#### Data definitions for table "lakes" 
* This table has the morphometry and elevation for the lmorpho lakes
* observations = `r l_n` 
* fields = `r l_f`

field	| source | units | indexed |	primary |	description
----------- | -----------	| ------- | ---- | ---- |	--------------------------------- 
**cellnum**|prism|NA|YES||the grid cell number of the prism raster(s) for the lake
**comid**|lmorpho|int|YES|primary|nhdplus comid from the lmorpho dataset
**longitude**|lmorpho|dd|||longitude of lmorpho centroid. Transformed to prism crs =="+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0"
**latitude**|lmorpho|dd|||latitude of lmorpho centroid. Transformed to prism crs =="+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0"
**surface_area**|lmorpho|m2|||lake surface area
**shoreline_length**|lmorpho|m |||length of lake shoreline
**shoreline_dev**|lmorpho|NA|||shoreline development index
**max_length**|lmorpho|m|||maximum length of the lake polygon
**max_width**|lmorpho|m|||maximum width of the lake polygon
**mean_width**|lmorpho|m|||mean width of the lake polygon
**max_depth**|lmorpho|m|||estimated maximum depth of lake
**mean_depth**|lmorpho|m|||estimated mean depth of lake
**volume**|lmorpho|m3|||estimated volume depth of lake
**elevation**|aws|m|||elevation of the lake

#### Data definitions for table "nla" 
* This table has the NLA water temperature and NLCD impervious data
* observations = `r n_n` 
* fields = `r n_f`

field	| source | units | indexed |	primary |	description
----------- | -----------	| ------- | ---- | ---- |	--------------------------------- 
**nla_id**|nla|char|YES|composite|unique identifiers for the NLA lakes
**comid**|lmorpho|int|YES||nhdplus comid from the lmorpho dataset
**cellnum**|prism|NA|||the grid cell number of the prism raster(s) for the lake
**year**|nla|YYYY|||sample year 2007 or 2012
**date**|nla|YYYYMMDD|||date the profile temperatures were collected
**visit_no**|nla|int||composite|All lakes were sampled once (visit_no = 1); some were sampled twice (visit_no = 2)
**tmean_2m**|nla|degrees C|||mean water temperature for depth <= 2 m
**tmean_2m_n**|nla|degrees C|||number of observations used to calculate tmean_2m
**percent_impervious**|nlcd|percent|||percent impervious surface in a 3000m buffer around the lake (includes islands); only observations with 100% coverage kept (i.e., lakes with partial NLCD coverage such as border lakes are excluded)
**area_impervious**|nlcd|m2|||m2 of impervious surface in a 3000m buffer around the lake (includes islands); only observations with 100% coverage kept (i.e., lakes with partial NLCD coverage such as border lakes are excluded)
**area_total**|lmorpho|m2|||area of 3000 m buffer around the lmorpho lake

#### Data definitions for table "tmeans" 
* This table has the aggregated prism air temperature estimates
* observations = `r tm_n` 
* fields = `r tm_f`

field	| source | units | indexed |	primary |	description
----------- | -----------	| ------- | ---- | ---- |	--------------------------------- 
**cellnum**|prism|NA|YES|composite|the grid cell number of the prism raster(s) for the lake
**date**|prism|YYYYMMDD|YES|composite|date of the prism estimate
**tmean**|prism|degrees C|||mean temperature on the date
**tmean_dm1**|prism|degrees C|||mean temperature on the day prior to the date (date minus 1)
**tmean_avg3**|prism|degrees C|||average of prism means for the 3 days prior to the date
**tmean_avg7**|prism|degrees C|||average of prism means for the 7 days prior to the date
**tmean_avg30**|prism|degrees C|||average of prism means for the 30 days prior to the date

#### Data definitions for table "tmeans_raw" 
* This table has the raw (by date) prism air temperature estimates
* observations = `r tmr_n` 
* fields = `r tmr_f`

field	| source | units | indexed |	primary |	description
----------- | -----------	| ------- | ---- | ---- |	--------------------------------- 
**cellnum**|prism|NA|YES|composite|the grid cell number of the prism raster(s) for the lake
**d0501:d0531**|prism|degrees C|||estimated mean temperature for dates May 1:May 31 by tmeans_raw.year
**d0601:d0630**|prism|degrees C|||estimated mean temperature for dates June 1:June 30 by tmeans_raw.year
**d0701:d0731**|prism|degrees C|||estimated mean temperature for dates July 1:July 31 by tmeans_raw.year
**d0801:d0831**|prism|degrees C|||estimated mean temperature for dates August 1:August 31 by tmeans_raw.year
**d0901:d0930**|prism|degrees C|||estimated mean temperature for dates September 1:September 30 by tmeans_raw.year
**year**|prism|YYYY|YES|composite|year 1981 to 2017

## Create a subset of the data for lake temperature model and save as nla_base.Rds

# Set up Data
* load nla data from "data_local/lake_temperatures.sqlite"
* create nla_base with the 2007 and 2012 nla, prism, lmorpho & imperv data for random forest analysis

```{r bdata, include=FALSE, echo=FALSE, cache=TRUE}
# load the nla data from lake_temperatures.sqlite

if(file.exists(here::here('data/nla_base.Rds'))) {
  nla_base <- readRDS(here::here('data/nla_base.Rds')) 
  } else {

means <- dbConnect(RSQLite::SQLite(), here::here("data_local/lake_temperatures.sqlite"))

# get the nla_base data
nla_base <- dbGetQuery(means, '
  SELECT nla.nla_id, nla.comid, lakes.cellnum, nla.year, nla.date, nla.visit_no, 
    nla.tmean_2m, nla.percent_impervious, lakes.longitude, lakes.latitude, lakes.surface_area, 
    lakes.shoreline_length, lakes.shoreline_dev, lakes.max_length, lakes.max_depth, lakes.max_width, 
    lakes.mean_width, lakes.elevation, lakes.mean_depth, lakes.volume
  FROM nla INNER JOIN lakes ON nla.comid = lakes.comid')

# add fields for tmeans
nla_base$tmean <- NA
nla_base$tmean_dm1 <- NA
nla_base$tmean_avg3 <- NA
nla_base$tmean_avg7 <- NA
nla_base$tmean_avg30 <- NA

for(i in min(which(is.na(nla_base$tmean))):nrow(nla_base)){
  print(i)
  print(Sys.time())
tm <- dbGetQuery(means, paste0("SELECT * FROM 'tmeans' WHERE cellnum = ", nla_base$cellnum[i], 
                         " AND date = ", nla_base$date[i]))

if(nrow(tm) > 0) {
  nla_base$tmean[i] <- tm$tmean
  nla_base$tmean_dm1[i] <- tm$tmean_dm1
  nla_base$tmean_avg3[i] <- tm$tmean_avg3
  nla_base$tmean_avg7[i] <- tm$tmean_avg7
  nla_base$tmean_avg30[i] <- tm$tmean_avg30
}}

dbDisconnect(means)

saveRDS(nla_base, file = here::here('data/nla_base.Rds')) # 2370   25
write.csv(nla_base, here::here('data/nla_base.csv'), row.names = FALSE)
}
```

### data definitions for nla_base.Rds & nla_base.csv

field	| source | units | indexed |	primary |	description
----------- | -----------	| ------- | ---- | ---- |	--------------------------------- 
**nla_id**|nla|char|YES|composite|unique identifiers for the NLA lakes
**comid**|lmorpho|int|YES||nhdplus comid from the lmorpho dataset
**cellnum**|prism|NA|||the grid cell number of the prism raster(s) for the lake
**year**|nla|YYYY|||sample year 2007 or 2012
**date**|nla|YYYYMMDD|||date the profile temperatures were collected
**visit_no**|nla|int||composite|All lakes were sampled once (visit_no = 1); some were sampled twice (visit_no = 2)
**tmean_2m**|nla|degrees C|||mean water temperature for depth <= 2 m
**percent_impervious**|nlcd|percent|||percent impervious surface in a 3000m buffer around the lake (includes islands); only observations with 100% coverage kept (i.e., lakes with partial NLCD coverage such as border lakes are excluded)
**longitude**|lmorpho|dd|||longitude of lmorpho centroid. Transformed to prism crs =="+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0"
**latitude**|lmorpho|dd|||latitude of lmorpho centroid. Transformed to prism crs =="+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0"
**surface_area**|lmorpho|m2|||lake surface area
**shoreline_length**|lmorpho|m |||length of lake shoreline
**shoreline_dev**|lmorpho|NA|||shoreline development index
**max_length**|lmorpho|m|||maximum length of the lake polygon
**max_depth**|lmorpho|m|||estimated maximum depth of lake
**max_width**|lmorpho|m|||maximum width of the lake polygon
**mean_width**|lmorpho|m|||mean width of the lake polygon
**elevation**|aws|m|||elevation of the lake
**mean_depth**|lmorpho|m|||estimated mean depth of lake
**volume**|lmorpho|m3|||estimated volume depth of lake
**tmean**|prism|degrees C|||mean temperature on the date
**tmean_dm1**|prism|degrees C|||mean temperature on the day prior to the date (date minus 1)
**tmean_avg3**|prism|degrees C|||average of prism means for the 3 days prior to the date
**tmean_avg7**|prism|degrees C|||average of prism means for the 7 days prior to the date
**tmean_avg30**|prism|degrees C|||average of prism means for the 30 days prior to the date