### 2019-04-30 R/model.Rmd
Change variable importance to MSE

---
title: "Untitled"
author: "Betty Kreakie"
date: "April 25, 2019"
output: html_document
---

```{r libraries, echo=TRUE, cache=TRUE}

library(randomForest)
library(dplyr)
library(purrr)
library(here)
library(broom)
library(ggplot2)
library(lubridate)
```

#Run Variable Selection stuff (hollister et al 2017)

```{r var_sel,echo=TRUE, cache=TRUE}

#varsel
source(here::here('R/varsel_regression_rf.R'))
source(here::here('R/varsel_plot.R'))

##Reading in and creating data sets
nla<-read.csv(here::here('data/nla_base.csv'))
x<-complete.cases(nla)
nla.cc<-nla[x,]
nla.cc$year<-factor(nla.cc$year)
nla.cc$date <- lubridate::yday(lubridate::ymd(nla.cc$date))

data07 <- filter(nla.cc, year == 2007, tmean_2m < 70) %>%
            dplyr::select(tmean_2m, date, longitude, latitude, tmean, tmean_dm1,
            tmean_avg3, tmean_avg7, tmean_avg30, percent_impervious, surface_area,
            shoreline_length, shoreline_dev, max_length, max_width, mean_width,
            max_depth, max_depth, volume, elevation)

data12 <- filter(nla.cc, year == 2012, tmean_2m < 70) %>%
            dplyr::select(tmean_2m, date, longitude, latitude, tmean, tmean_dm1,
            tmean_avg3, tmean_avg7, tmean_avg30, percent_impervious, surface_area,
            shoreline_length, shoreline_dev, max_length, max_width, mean_width,
            max_depth, max_depth, volume, elevation)

nla.all <-  filter(nla.cc, tmean_2m < 70) %>%
  dplyr::select(tmean_2m, year,date, longitude, latitude, tmean,          tmean_avg30,percent_impervious,surface_area,shoreline_length, shoreline_dev, max_length, max_width, mean_width, max_depth, max_depth, volume, elevation)

##variable selection for 2007-only model
var_sel_tmean07 <- varsel_regression_rf(data07$tmean_2m, select(data07, -tmean_2m), ntree=1000, prog=TRUE, importance=TRUE)

varsel_plot(var_sel_tmean07)

var_sel_tmean07$vars[[9]]

##Variable selection for 2012-only model
var_sel_tmean12 <- varsel_regression_rf(data12$tmean_2m, select(data12, -tmean_2m),                ntree=1000, prog=TRUE,importance=TRUE)

varsel_plot(var_sel_tmean12) #9 variables

var_sel_tmean12$vars[[9]]

##Variable selection for 2007&2012 model
var_sel_tmean <- varsel_regression_rf(nla.all$tmean_2m, select(nla.all, -tmean_2m),                ntree=1000, prog=TRUE, importance=TRUE)

varsel_plot(var_sel_tmean) #7 variables

var_sel_tmean$vars[[7]]

```

# Build our draft model with selected variables

```{r draft_model,echo=TRUE, cache=TRUE}

# Draft Model

# Using tmean made most sense and perfomed better that tmin and tmax based models.

RF2007 <- data07 %>%

  select(tmean_2m,var_sel_tmean07$vars[[9]]) %>%

  randomForest(tmean_2m ~ ., data = ., importance = TRUE,

               ntree = 10000)

RF2007

varImpPlot(RF2007)

RF2012 <- data12 %>%

  select(tmean_2m,var_sel_tmean12$vars[[9]]) %>%

  randomForest(tmean_2m ~ ., data = ., importance = TRUE,

               ntree = 10000)

RF2012

varImpPlot(RF2012)

RFAll <- nla.all %>%

  select(tmean_2m,var_sel_tmean$vars[[7]]) %>%

  randomForest(tmean_2m ~ ., data = ., importance = TRUE,

               ntree = 10000)

RFAll

varImpPlot(RFAll)

```

mean_run <- nla.cc %>%

  select(tmean_2m,var_sel_tmean$vars[[9]]) %>%

  randomForest(tmean_2m ~ ., data = ., importance = TRUE,

               ntree = 1000)

mean_run

varImpPlot(mean_run)

```